---
title: "Analyze Phase"
author: "Fisnik"
date: "2024-01-05"
output: html_document
---
## Average Data by users
```{r}
library(dplyr)

av_data_users <- Daily_Activity_Sleep %>%
  group_by(Id) %>%
  summarize(
    Avg_Distance = round(mean(TotalDistance, na.rm = TRUE), 2),
    Avg_Daily_Steps = round(mean(TotalSteps, na.rm = TRUE), 2),
    Avg_Sleep = round(mean(TotalMinutesAsleep, na.rm = TRUE), 0),
    Avg_Calories = round(mean(Calories, na.rm = TRUE), 0)
  )
 View(av_data_users)
```
## Average Data by users
```{r}
av_data_DayfWeek <- Daily_Activity_Sleep %>%
  group_by(Day_of_Week) %>%
  summarize(
    Avg_Distance = round(mean(TotalDistance, na.rm = TRUE), 2),
    Avg_Daily_Steps = round(mean(TotalSteps, na.rm = TRUE), 2),
    Avg_Sleep = round(mean(TotalMinutesAsleep, na.rm = TRUE), 0),
    Avg_Calories = round(mean(Calories, na.rm = TRUE), 0)
  )
View(av_data_DayfWeek)
```
 #Total Data by Users
```{r}
total_data_users <- Daily_Activity_Sleep %>%
  group_by(Id) %>%
  summarize(
    Total_Distance = round(sum(TotalDistance, na.rm = TRUE), 2),
    Total_Daily_Steps = sum(mean(TotalSteps, na.rm = TRUE), 2),
    Total_Sleep = round(sum(TotalMinutesAsleep, na.rm = TRUE), 0),
    Total_Calories = round(sum(Calories, na.rm = TRUE), 0)
  )

View(total_data_users)
```

#Total Data by Day of Week
```{r}
total_data_DayfWeek <- Daily_Activity_Sleep %>%
  group_by(Day_of_Week) %>%
  summarize(
    Total_Distance = round(sum(TotalDistance, na.rm = TRUE), 2),
    Toal_Daily_Steps = round(sum(TotalSteps, na.rm = TRUE), 2),
    Total_Sleep = round(sum(TotalMinutesAsleep, na.rm = TRUE), 0),
    Total_Calories = round(sum(Calories, na.rm = TRUE), 0)
  )
View(total_data_DayfWeek)
```
#New DataFrame called User Classification by AVG Daily Steps
```{r}
user_cat_by_avg_dailysteps <- av_data_users %>%
  mutate(
    User_Type = case_when(
      Avg_Daily_Steps < 5000 ~ "Sedentary",
      Avg_Daily_Steps >= 5000 & Avg_Daily_Steps < 7499 ~ "Lightly Active",
      Avg_Daily_Steps >= 7500 & Avg_Daily_Steps < 9999 ~ "Fairly Active",
      Avg_Daily_Steps >= 10000 ~ "Very Active",
      TRUE ~ NA_character_  # Default case if none of the above conditions are met
    )
  )

View(user_cat_by_avg_dailysteps)
```

#Determining the percentage of different types of users based on average daily steps
```{r}
# Step 1: Count the number of occurrences of each User_Type
user_counts <- user_cat_by_avg_dailysteps %>%
  count(User_Type)

# Step 2: Calculate the total number of users
total_users <- nrow(user_cat_by_avg_dailysteps)

# Step 3: Calculate the percentage for each User_Type
user_counts <- user_counts %>%
  mutate(Percentage = n / total_users * 100)

# View the results
user_counts
View(user_counts)
```
#Creating a pie chart for User Types
```{r}
# Creating the pie chart
# Prepare labels with percentages
labels_with_percentage <- paste(user_counts$User_Type, 
                                " - ", 
                                round(user_counts$Percentage, 2), "%", sep = "")

# Creating the pie chart with percentage labels
pie(user_counts$Percentage, 
    labels = labels_with_percentage, 
    main = "User Classification by Average Daily Steps", 
    col = rainbow(length(user_counts$User_Type)))


```
#Average Hourly Steps Throughout the Day
```{r}
average_hourly_steps <- hourly_steps %>%
  group_by(Time) %>%
  summarize(Avg_StepTotal = round(mean(StepTotal, na.rm = TRUE), 0))

View(average_hourly_steps) 

```

#Plotting the Bar Chart Using ggplot2
```{r}
# Load the ggplot2 library
library(ggplot2)

# Create the bar chart with a color gradient
ggplot(average_hourly_steps, aes(x = Time, y = Avg_StepTotal, fill = Avg_StepTotal)) +
  geom_bar(stat = "identity", width = 0.7) +  # Use stat = "identity" to use the y values for bar heights
  scale_fill_gradient(low = "navy", high = "red") +  # Color gradient from navy to red
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +  # Rotate the x-axis text for readability
  labs(x = "Time", y = "Average Steps", title = "Average Hourly Steps") +
  scale_x_discrete(limits = average_hourly_steps$Time)  # Ensure the x-axis has all the hours in order

```
#Calculating Average Calories Throughout the Day
```{r}
average_hourly_calories <- hourly_calories %>%
  group_by(Time) %>%
  summarize(Avg_Calories = round(mean(Calories, na.rm = TRUE), 0))

View(average_hourly_calories)
```

#Plotting the barchart using ggplot2
```{r}
# Load necessary libraries
library(ggplot2)
library(scales)  # For the gradient color scale
# Create the bar chart
ggplot(average_hourly_calories, aes(x = Time, y = Avg_Calories)) +
  geom_bar(stat = "identity", aes(fill = Avg_Calories), width = 0.7) +  # Use fill to add the gradient
  geom_text(aes(label = Avg_Calories), vjust = -0.3, size = 2.5) +  # Add labels on top of the bars
  scale_fill_gradient2(low = "navy", high = "red", mid = "yellow", midpoint = median(average_hourly_calories$Avg_Calories)) +  # Gradient color scale
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +  # Rotate x-axis labels for readability
  labs(x = "Time of Day", y = "Average Calories Burned", title = "Average Hourly Calories Burned")

```
#Calculating Usage in Days by Users

```{r}
library(dplyr)
# Creating the new dataframe with Id and Total_Usage_in_Days
frequency_of_usage <- Daily_Activity_Sleep %>%
  group_by(Id) %>%
  summarize(Total_Usage_in_Days = n())

View(frequency_of_usage)
```

#Classification of Users by use of the app in the number of days
```{r}
# Assuming frequency_of_usage already exists with the columns Id and Total_Usage_in_Days
classification_usage_days <- frequency_of_usage %>%
  mutate(Type_of_Usage = case_when(
    Total_Usage_in_Days >= 1 & Total_Usage_in_Days <= 10 ~ "Low Use",
    Total_Usage_in_Days >= 11 & Total_Usage_in_Days <= 20 ~ "Moderate Use",
    Total_Usage_in_Days >= 21 & Total_Usage_in_Days <= 31 ~ "High Use",
    TRUE ~ "Other"  # For cases that don't match any of the above conditions
  ))

View(classification_usage_days)
```

#Calculating the Percentage of users based on use of the app in days
```{r}
# Calculate the percentage of each Type_of_Usage
usage_percentages <- classification_usage_days %>%
  count(Type_of_Usage) %>%
  mutate(Percentage = n / 24 * 100)

View(usage_percentages)
```
#Creating a Pie Chart for the Usage Type
```{r}
# Prepare the labels with percentage
library(ggplot2)
usage_percentages$labels <- paste(usage_percentages$Type_of_Usage, round(usage_percentages$Percentage, 1), "%", sep=" ")
# Create the pie chart with specific colors
ggplot(usage_percentages, aes(x = "", y = Percentage, fill = Type_of_Usage)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = c("navy", "yellow", "red")) + # Manually assign colors
  geom_text(aes(label = labels), position = position_stack(vjust = 0.5)) +
  theme_void() +
  labs(title = "Percentage of Each Type of Usage")
```
```{r}
# Load the necessary libraries
library(ggplot2)
library(ggrepel)

# Assuming usage_percentages is your dataframe and it already contains a 'labels' column 
# Create the pie chart with labels outside the slices
ggplot(usage_percentages, aes(x = "", y = Percentage, fill = Type_of_Usage)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  scale_fill_manual(values = c("navy", "yellow", "red")) + # Manually set the colors
  geom_label_repel(aes(label = labels), 
                   nudge_x = 0.5, nudge_y = 0.5, # Adjust the nudge values as needed
                   size = 3,
                   box.padding = unit(0.35, "lines"),
                   point.padding = unit(0.5, "lines")) +
  theme_void() +
  labs(title = "Percentage of Each Type of Usage") +
  theme(legend.position = "right")

```
#Total Time spent on Each Activity per Day
```{r}
daily_activity_summary <- daily_activity %>%
  group_by(Day_of_Week) %>%
  summarize(
    Sedentary_Mins = sum(SedentaryMinutes, na.rm = TRUE),
    LightlyActive_Mins = sum(LightlyActiveMinutes, na.rm = TRUE),
    FairlyActive_Mins = sum(FairlyActiveMinutes, na.rm = TRUE),
    VeryActive_Mins = sum(VeryActiveMinutes, na.rm = TRUE)
  )
View(daily_activity_summary)
```

#Total Time spent on Each Activity per Day
```{r}
# Load necessary libraries
library(ggplot2)
library(tidyr)
library(dplyr)

# Create a dataframe from your provided data
daily_activity_summary <- data.frame(
  Day_of_Week = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"),
  Sedentary_Mins = c(123353, 153119, 148422, 141413, 126039, 119571, 119821),
  LightlyActive_Mins = c(23047, 29996, 28478, 27257, 25729, 25686, 21051),
  FairlyActive_Mins = c(1680, 2179, 1965, 1758, 1526, 1885, 1758),
  VeryActive_Mins = c(2773, 3489, 3117, 2853, 2527, 2718, 2418)
)

# Ensure Day_of_Week is a factor and in the correct order
daily_activity_summary$Day_of_Week <- factor(daily_activity_summary$Day_of_Week, 
                                             levels = c("Monday", "Tuesday", "Wednesday", "Thursday", 
                                                        "Friday", "Saturday", "Sunday"))

# Reshape data from wide to long format
daily_activity_long <- pivot_longer(daily_activity_summary, 
                                    cols = -Day_of_Week, 
                                    names_to = "ActivityType", 
                                    values_to = "Minutes")

# Create the grouped bar chart
ggplot(daily_activity_long, aes(x = Day_of_Week, y = Minutes, fill = ActivityType)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.75)) +
  scale_fill_manual(values = c("Sedentary_Mins" = "navy", 
                               "LightlyActive_Mins" = "yellow", 
                               "FairlyActive_Mins" = "orange", 
                               "VeryActive_Mins" = "red")) +
  labs(x = "Day of the Week", y = "Total Minutes", 
       title = "Activity Minutes by Day of the Week") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1))  # Adjust text angle for readability



```
##Total Minutes each user spends on each activity
```{r}
library(dplyr)
# Assuming your dataframe is named daily_activity
total_activity_by_user <- daily_activity %>%
  group_by(Id) %>%
  summarize(
    Sedentary_Mins = sum(SedentaryMinutes, na.rm = TRUE),
    LightlyActive_Mins = sum(LightlyActiveMinutes, na.rm = TRUE),
    FairlyActive_Mins = sum(FairlyActiveMinutes, na.rm = TRUE),
    VeryActive_Mins = sum(VeryActiveMinutes, na.rm = TRUE)
  )

View(total_activity_by_user)
```

##Calculating Percentage of Total Time spent on Each Activity per Day

```{r}
# Calculate total minutes for each day
daily_activity_summary$Total_Mins <- rowSums(daily_activity_summary[c("Sedentary_Mins", "LightlyActive_Mins", "FairlyActive_Mins", "VeryActive_Mins")])

# Calculate percentages
daily_activity_summary$Sedentary_Percent <- daily_activity_summary$Sedentary_Mins / daily_activity_summary$Total_Mins * 100
daily_activity_summary$LightlyActive_Percent <- daily_activity_summary$LightlyActive_Mins / daily_activity_summary$Total_Mins * 100
daily_activity_summary$FairlyActive_Percent <- daily_activity_summary$FairlyActive_Mins / daily_activity_summary$Total_Mins * 100
daily_activity_summary$VeryActive_Percent <- daily_activity_summary$VeryActive_Mins / daily_activity_summary$Total_Mins * 100

# Select only the relevant columns to display
percentage_data <- daily_activity_summary[c("Day_of_Week", "Sedentary_Percent", "LightlyActive_Percent", "FairlyActive_Percent", "VeryActive_Percent")]
percentage_data

# Round the percentages to two decimal places and format as strings with a "%" sign
percentage_data$Sedentary_Percent <- sprintf("%.2f%%", round(percentage_data$Sedentary_Percent, 2))
percentage_data$LightlyActive_Percent <- sprintf("%.2f%%", round(percentage_data$LightlyActive_Percent, 2))
percentage_data$FairlyActive_Percent <- sprintf("%.2f%%", round(percentage_data$FairlyActive_Percent, 2))
percentage_data$VeryActive_Percent <- sprintf("%.2f%%", round(percentage_data$VeryActive_Percent, 2))

# Display the updated dataframe
View(percentage_data)

```

### Joining the dataframes and creating Daily_Use
```{r}
# Joining the dataframes and creating Daily_Use
Daily_Use <- left_join(daily_activity, classification_usage_days[, c("Id", "Total_Usage_in_Days", "Type_of_Usage")], by = "Id")

View(Daily_Use)

# Add Total_Minutes_Worn column to Daily_Use
Daily_Use <- Daily_Use %>%
  mutate(Total_Minutes_Worn = VeryActiveMinutes + FairlyActiveMinutes + LightlyActiveMinutes + SedentaryMinutes)

# Load dplyr library
library(dplyr)

# Add Worn_Type column to Daily_Use
Daily_Use <- Daily_Use %>%
  mutate(Worn_Type = case_when(
    (Total_Minutes_Worn / 1440) * 100 == 100 ~ "All Day",
    (Total_Minutes_Worn / 1440) * 100 < 100 & (Total_Minutes_Worn / 1440) * 100 >= 50 ~ "More Than Half Day",
    (Total_Minutes_Worn / 1440) * 100 < 50 & (Total_Minutes_Worn / 1440) * 100 > 0 ~ "Less Than Half Day",
    TRUE ~ NA_character_  # For cases that don't match any of the above conditions or missing values
  ))
```


```{r}
# Remove rows with NA values in any column from Daily_Use
Daily_Use_Cleaned <- na.omit(Daily_Use)
View(Daily_Use_Cleaned)
```

```{r}
library(dplyr)

# Assuming Daily_Use_Cleaned is your dataframe and it contains the column Worn_Type
worn_type_summary <- Daily_Use_Cleaned %>%
  group_by(Worn_Type) %>%
  summarize(
    Total_Users = n(),
    Worn_Type_Percentage = paste0(round((n() / 713) * 100, 1), "%")
  )

View(worn_type_summary)
```

```{r}
# Convert the Worn_Type_Percentage from a character to a numeric format
worn_type_summary$Numeric_Percentage <- as.numeric(sub("%", "", worn_type_summary$Worn_Type_Percentage))

# Create the pie chart
ggplot(worn_type_summary, aes(x = "", y = Numeric_Percentage, fill = Worn_Type)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  theme_void() +
  scale_fill_brewer(palette = "Set3") +
  geom_text(aes(label = Worn_Type_Percentage), position = position_stack(vjust = 0.5))
```
##Correlation Between Daily Steps and Minutes Asleep
```{r}
stepsminutes_correlation <- Daily_Activity_Sleep %>%
  select(Steps = TotalSteps, TotalMinutesAsleep)

View(stepsminutes_correlation)
```

## Use Scatterplot in ggplot2 to visualize correlation between Daily Steps and Minutes Asleep
```{r}
# Load necessary library
library(ggplot2)

# Create the scatter plot with a curved trend line
ggplot(stepsminutes_correlation, aes(x = Steps, y = TotalMinutesAsleep)) +
  geom_point() +  # Add points
  geom_smooth(method = "loess", span = 1, color = "blue", se = FALSE) +  # Add curved trend line
  labs(x = "Daily Steps", y = "Minutes Asleep", 
       title = "Correlation between Daily Steps and Minutes Asleep") +
  theme_minimal()  # Use a minimal theme


```

##Correlation Between Daily Steps and Calories
```{r}
stepscalories_correlation <- Daily_Activity_Sleep %>%
  select(Steps = TotalSteps, Calories)

View(stepscalories_correlation)
```
## Use Scatterplot in ggplot2 to visualize correlation between Daily Steps and Calories

```{r}
# Load necessary library
library(ggplot2)

# Create the scatter plot with a curved trend line
ggplot(stepscalories_correlation, aes(x = Steps, y = Calories)) +
  geom_point() +  # Add points
  geom_smooth(method = "loess", span = 0.9, color = "blue", se = FALSE) +  # Add curved trend line
  labs(x = "Daily Steps", y = "Calories", 
       title = "Correlation between Daily Steps and Calories") +
  theme_minimal()  # Use a minimal theme

```










