---
title: "Process and Analysis BB"
author: "Fisnik"
date: "2023-12-29"
output: html_document
---
Processing the Data

# Load the dplyr package
```{r}
library(dplyr)
```
Assuming dailyactivity and sleepday data frames are already loaded in your R session
SQL Query 1: SELECT count(DISTINCT(Id)) FROM daily_activity
```{r}
distinct_count_dailyactivity <- daily_activity %>%
  summarise(count_distinct_id = n_distinct(Id))
```
SQL Query 2: SELECT count(DISTINCT(Id)) FROM sleep_day
```{r}
distinct_count_sleepday <- sleep_day %>%
  summarise(count_distinct_id = n_distinct(Id))
```
# Display the results
```{r}
distinct_count_dailyactivity
distinct_count_sleepday
```

## Distinct Values in Hourly Steps and Hourly Calories
```{r}
n_distinct(hourly_steps$Id)
n_distinct((hourly_calories$Id))
```

Checking for Duplicates on Daily Activity
```{r}
# Find duplicates based on specific columns, e.g., 'Id' and 'ActivityDate'
duplicates_da <- daily_activity[duplicated(daily_activity[c("Id", "ActivityDate", "TotalSteps")]), ]

# View the duplicate rows
head(duplicates_da)
```
Checking for Duplicates on Sleep Day
```{r}
# Find duplicates based on specific columns, e.g., 'Id' and 'ActivityDate'
duplicates_sd <- sleep_day[duplicated(sleep_day[c("Id", "SleepDay", "TotalSleepRecords")]), ]

# View the duplicate rows
head(duplicates_sd)
```
# Count the number of duplicate rows
```{r}
num_duplicates <- sum(duplicated(sleep_day))
```

# Print the result
```{r}
num_duplicates
```

Checking for Duplicates on Hourly Steps
```{r}
# Find duplicates based on specific columns, e.g., 'Id' and 'ActivityDate'
duplicates_hs <- hourly_steps[duplicated(hourly_steps[c("Id", "Date", "Time", "StepTotal")]), ]

# View the duplicate rows
head(duplicates_hs)
```
Checking for Duplicates on Hourly Calories
```{r}
# Find duplicates based on specific columns, e.g., 'Id' and 'ActivityDate'
duplicates_hc <- hourly_calories[duplicated(hourly_calories[c("Id", "Date", "Time", "Calories")]), ]

# View the duplicate rows
head(duplicates_hc)
```
##Deleting duplicate records from sleepday
```{r}
# Remove duplicate rows
sleep_day <- sleep_day %>% 
  distinct()

# View the updated data frame
head(sleep_day)

```
Upload Lubridate and Dplyr
```{r}
library(lubridate)
library(dplyr)
```

#(hourly_steps) Split the ActivityHour column into separate date and time columns and format the date:
```{r}
#Splitting and Formatting
hourly_steps <- hourly_steps %>%
  mutate(
    Date = format(as.Date(mdy_hms(ActivityHour)), "%m/%d/%Y"),
    Time = format(mdy_hms(ActivityHour), "%H:%M:%S")
  )
## Remove the ActivityHour column from hourly_steps
hourly_steps <- hourly_steps %>%
  select(-ActivityHour)

# Reorder columns in hourly_steps
hourly_steps <- hourly_steps %>%
  select(Id, Date, Time, StepTotal)

```

#(hourly_calories) Split the ActivityHour column into separate date and time columns and format the date:
```{r}
#Splitting and Formatting
hourly_calories <- hourly_calories %>%
  mutate(
    Date = format(as.Date(mdy_hms(ActivityHour)), "%m/%d/%Y"),
    Time = format(mdy_hms(ActivityHour), "%H:%M:%S")
  )
# Remove the ActivityHour column from hourly_calories
hourly_calories <- hourly_calories %>%
  select(-ActivityHour)
# Reorder columns in hourly_calories
hourly_calories <- hourly_calories %>%
  select(Id, Date, Time, Calories)
```

#Formatting the date on Sleep Day table:
```{r}
# Assuming sleep_day is your data frame

# 1. Convert SleepDay to POSIXct
sleep_day$SleepDay <- as.POSIXct(sleep_day$SleepDay, format="%m/%d/%Y %I:%M:%S %p", tz="UTC")

# 2. Extract Date and Time
sleep_day$Date <- as.Date(sleep_day$SleepDay) # Extract Date
sleep_day$Time <- format(sleep_day$SleepDay, "%H:%M:%S") # Extract Time

# 3. Remove the SleepDay column
sleep_day <- sleep_day[, !colnames(sleep_day) %in% "SleepDay"]
 
# 1. Reorder Columns to Place Date as the Second Column
sleep_day <- sleep_day[, c("Id", "Date", setdiff(names(sleep_day), c("Id", "Date", "Time")))]

# Remove the Time column from sleep_day
sleep_day$Time <- NULL

# Assuming sleep_day is your data frame and 'Id' is the first column

# Rearrange so that Date is the second column
sleep_day <- sleep_day[, c("Id", "Date", setdiff(names(sleep_day), c("Id", "Date")))]

# Convert the Date column to the new format
sleep_day$Date <- format(as.Date(sleep_day$Date), "%m/%d/%Y")

# Rename the column using dplyr
sleep_day <- sleep_day %>% rename(SleepDay = Date)
```


```{r}
# Convert ActivityDate from character to Date
# Replace "%Y-%m-%d" with the format that matches your date strings
daily_activity$ActivityDate <- as.Date(daily_activity$ActivityDate, format = "%m/%d/%Y")

# Add Day_of_Week column
daily_activity$Day_of_Week <- weekdays(daily_activity$ActivityDate)

```

##Update all the tables date format since in R date is formated to "%Y-%m-%d"
```{r}
hourly_steps$Date <- as.Date(hourly_steps$Date, format = "%m/%d/%Y")
hourly_calories$Date <- as.Date(hourly_calories$Date, format = "%m/%d/%Y")

```

